---
title: "PerDayTrendSection"
author: "Mildred Ouyang"
date: "2022-12-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(plyr)
library(dplyr)
library(tidyverse)

#install.packages("hash")
library(hash)
#install.packages("reticulate")
library(reticulate)
```

```{r}
process_name <- function(csv){
  return(gsub("  ", "", gsub(".csv", "", gsub("_", " ", csv))))
}

get_attraction_name <- function(){
  return(lapply(list.files("/Users/mildredouyang/Desktop/5702/EDAV-Final-Project/Data/data/Magic Kingdom/xysong_python"), process_name))
}

filenames <- list.files("/Users/mildredouyang/Desktop/5702/EDAV-Final-Project/Data/data/Magic Kingdom/xysong_python")
dict <- hash()
for (csv in filenames) {
  dict[[process_name(csv)]] <- read.csv(sprintf("/Users/mildredouyang/Desktop/5702/EDAV-Final-Project/Data/data/Magic Kingdom/xysong_python/%s", csv)) 
}
```

```{r}
Adventureland <- c("enchanted tiki rm", "jungle cruise", "magic carpets", "pirates of caribbean")

Fantasyland <- c("7 dwarfs train", "dumbo", "it s a small world", "peter pan s flight", "winnie the pooh", "under the sea", "regal carrousel", "mad tea party", "barnstormer")

Frontierland <- c("big thunder mtn", "splash mountain", "country bears")

Tomorrowland <- c("buzz lightyear", "astro orbiter", "space mountain", "tom land speedway", "peoplemover", "carousel of progress")

Liberty_Square <- c("haunted mansion", "hall of presidents", "liberty sq riverboat")

section_name <- list("Adventureland", "Fantasyland", "Frontierland", "Tomorrowland", "Liberty Square")
```

```{r}
adventure <- bind_rows(lapply(Adventureland, function(i) dict[[i]]))
```

```{r}
fantasy <- bind_rows(lapply(Fantasyland, function(i) dict[[i]]))
```

```{r}
frontier <- bind_rows(lapply(Frontierland, function(i) dict[[i]]))
```

```{r}
tomorrow <- bind_rows(lapply(Tomorrowland, function(i) dict[[i]]))
```

```{r}
liberty <- bind_rows(lapply(Liberty_Square, function(i) dict[[i]]))
```

```{r}
avg_section <- c(mean(adventure$Waiting.Time), mean(fantasy$Waiting.Time), mean(frontier$Waiting.Time), mean(tomorrow$Waiting.Time), mean(liberty$Waiting.Time))
```

```{r}
section_avg = data.frame(unlist(section_name), unlist(avg_section))
names(section_avg) = c("section","avg")
```

```{r}
ggplot(section_avg, aes(x=avg, y=fct_reorder(section, avg))) + 
  geom_bar(stat = "identity", fill = "#393E8F") +
  labs(title = "Average Waiting Time for Each Land", 
       x = "Average Waiting Time", 
       y = "Land")
```

The above bar chart shows the average wait time by each themed land, and it is sorted by descending order from the top. Frontierland has the longest average wait time followed by Tomorrowland. The rest have similar average wait time.       
However, this chart illustrated the Simpson's Paradox. The top 3 longest average wait time attractions - Seven Dwarfs Mine Train, Space Mountain, and Peter Pan's Flight - are not in the Frontierland.    
```{r}
adventure$Datetime <- as.POSIXlt(adventure$Datetime)
fantasy$Datetime <- as.POSIXlt(fantasy$Datetime)
frontier$Datetime <- as.POSIXlt(frontier$Datetime)
tomorrow$Datetime <- as.POSIXlt(tomorrow$Datetime)
liberty$Datetime <- as.POSIXlt(liberty$Datetime)
```

```{r}
m_savg <- list()
a_savg <- list()
e_savg <- list()

section_name <- list("Adventureland", "Fantasyland", "Frontierland", "Tomorrowland", "Liberty Square")

m_savg <- append(m_savg, mean(adventure[adventure$Datetime$hour < 12, 'Waiting.Time']))
m_savg <- append(m_savg, mean(fantasy[fantasy$Datetime$hour < 12, 'Waiting.Time']))
m_savg <- append(m_savg, mean(frontier[frontier$Datetime$hour < 12, 'Waiting.Time']))
m_savg <- append(m_savg, mean(tomorrow[tomorrow$Datetime$hour < 12, 'Waiting.Time']))
m_savg <- append(m_savg, mean(liberty[liberty$Datetime$hour < 12, 'Waiting.Time']))

a_savg <- append(a_savg, mean(adventure[((adventure$Datetime$hour >= 12) | (adventure$Datetime$hour < 18)), 'Waiting.Time']))
a_savg <- append(a_savg, mean(fantasy[((fantasy$Datetime$hour >= 12) | (fantasy$Datetime$hour < 18)), 'Waiting.Time']))
a_savg <- append(a_savg, mean(frontier[((frontier$Datetime$hour >= 12) | (frontier$Datetime$hour < 18)), 'Waiting.Time']))
a_savg <- append(a_savg, mean(tomorrow[((tomorrow$Datetime$hour >= 12) | (tomorrow$Datetime$hour < 18)), 'Waiting.Time']))
a_savg <- append(a_savg, mean(liberty[((liberty$Datetime$hour >= 12) | (liberty$Datetime$hour < 18)), 'Waiting.Time']))

e_savg <- append(e_savg, mean(adventure[adventure$Datetime$hour >= 18, 'Waiting.Time']))
e_savg <- append(e_savg, mean(fantasy[fantasy$Datetime$hour >= 18, 'Waiting.Time']))
e_savg <- append(e_savg, mean(frontier[frontier$Datetime$hour >= 18, 'Waiting.Time']))
e_savg <- append(e_savg, mean(tomorrow[tomorrow$Datetime$hour >= 18, 'Waiting.Time']))
e_savg <- append(e_savg, mean(liberty[liberty$Datetime$hour >= 18, 'Waiting.Time']))

m_savgtime = data.frame(unlist(as.character(section_name)), unlist(m_savg))
names(m_savgtime) = c("section","avg")
a_savgtime = data.frame(unlist(as.character(section_name)), unlist(a_savg))
names(a_savgtime) = c("section","avg")
e_savgtime = data.frame(unlist(as.character(section_name)), unlist(e_savg))
names(e_savgtime) = c("section","avg")
```

```{r}
m_savgtime['timeofday'] <- 'morning'
a_savgtime['timeofday'] <- 'afternoon'
e_savgtime['timeofday'] <- 'evening'
savg_comb <- NA
savg_comb <- rbind(savg_comb, m_savgtime)
savg_comb <- rbind(savg_comb, a_savgtime)
savg_comb <- rbind(savg_comb, e_savgtime)
savg_comb <- na.omit(savg_comb)
```

```{r}
ggplot(savg_comb, aes(x=avg, y=fct_reorder2(section,timeofday=='afternoon',avg,.desc=FALSE), color = timeofday)) + 
  geom_point() +
  labs(title = "Average Waiting Time for Each Land", 
       x = "Average Waiting Time", 
       y = "Land") +
  guides(color = guide_legend(title = "Time of Day"))
```

The above Cleveland dot plot shows the average wait time by each land at different time of the day. The plot is sorted by the average afternoon wait time in descending order from top to bottom. The order of themed lands is the same as the one in the average wait time plot. For all themed lands, the average afternoon wait time is the longest. The average evening time is the shortest for 4 of the 5 themed lands. Only one themed land has morning as the the shortest average wait time.    